---
title: "Arch 241 example"
author: "Dana"
date: "August 18, 2018"
output: html_document
---
# Introduction

Here is a "chunk" of R code. You can execute it one line at a time, or all at once. 

```{r}
  x<-9 
  sqrt(x)
  log(x)  # You can write notes
```

You can also write text between your chunks of code

# First operations in R

## Creating a list and selecting elements from a list
```{r}

  x <- c(3,-2,4,2,0,6)
  x
  x[3]
  x[1:3]
  mean(x)
  sum(x)
  max(x)
  
  z <- x > 2; z
```
  
  
## Assigning some more variables
```{r}
  v1 = seq(1,10, by=0.5) #Variables are case sensitive
  V1 = 'Ciao'
  v1;V1
  
  sd(v1)
  
  v2  <-v1^2
  v3 <- v1[v1>5];v3
```
  
## Making basic plots with base r
```{r}
  plot(v1,v2, type='p')
  help(plot)
  plot(v1,v2, type='l')
  plot(v1,v2, type='b')
```
  
```{r}
plot(sin(seq(0,8*pi,length=100)),type='l')
summary(v1)
plot(v1,v2, type='l',ylab='V2', xlab='V1',main='Arch 241', sub='First plot')
```

## Running a linear model on the two plotted variables
```{r}
  cor(v1,v2)
  LinearModel<-lm(v2~v1); summary(LinearModel)
```       

# Working with thermal comfort data
### load some helpful packages
```{r}
library(tidyverse)
library(here)
library(skimr) # a summary function
```

### Load file
```{r}
Arch241simple <- read_csv(here("example_data.csv"))
Arch241simple
```

### summary
```{r}
# 
Arch241simple[1:30,] # the first 30 columns
summary(Arch241simple) #stat summary
skim(Arch241simple)

#summary(Arch241simple$Top) # not working as expected

names(Arch241simple) #variable names
```

### Subsetting
```{r}
#tidyverse style
Arch241simpleMale <- Arch241simple %>%
    filter(Sex=="M")

summary(Arch241simpleMale) 

#Older base R style
#Arch241simpleMale <- subset(Arch241simple, #Gender=="Male");summary(Arch241simpleMale) #subset
```
